Nilpotent Invariants \\ for Generic Discrete Series of Real Groups

### 2024-09-10 whittaker.tex
some proofreading

We need to pass to the dual Lie algebra $\g^*$.
We say an element of $\g^*$ is nilpotent if its $G$-orbit is a (weak) cone, i.e. closed under multiplication by $\C^*$,
and \emph{regular} if $\dim_{\Cent(G)}(X)=\mathrm{rank}(G)$.  
The principal nilpotent orbit $\O_p\in \g^*$ is the set of regular nilpotent elements.

The statements of the main results do not require the choice of an isomorphism $\g\simeq \g^*$.
In several places we do need to make such a choice;
with the exception of (?) this choice will not matter.

So we fix a non-degenerate, $G(\R)$-invariant bilinear form $\kappa$ on $\g(\R)$. This is equivalent to having a non-degenerate, $G(\C)$ and Galois invariant
bilinear form on $\g$. The restriction of the form each simple factor
is unique up to a non-zero constant. For example we can use the Killing form on $\gder$, extended arbitrarily by a non-degenerate form on the center.

@@

Using $\psi_\kappa$ we transfer the asymptotic cone, Kostant section, and Kostant Sekiguchi corresondence to $\g^*$. In each case these constructions are independent of the
choice of $\kappa$. See (?).

In Section \ref{s:AV} we need to be more precise: we choose  $\kappa$ to be the Killing form. (More generally, we could take any non-degenerate
invariant form which is negative definite on the Lie algebra of a maximal compact subgroup.)

@@

It is well known that a natural bijection between nilpotent $G(\R)$-orbits on $\g(\R)$ and nilpotent $K$ orbits on $\s$. 
This is the Kostant--Sekiguchi
correspondence~\cite{sekiguchi}, which we write as $\KS(\O)$. It is a deep theorem of
Schmid and Vilonen (the main result of~\cite{SV1}) that the wavefront
set and the associated variety are exchanged by this
correspondence. (As a historical aside, the Kostant--Sekiguchi
correspondence was discovered on the basis that the correspondence
$\AV(\pi) \leftrightarrow \WF(\pi)$ might be expressed in elementary
terms.)

@@

\begin{proposition} \label{Sekiguchi_result}
Suppose $\pi$ is a generic discrete series representation. Then
$$
\KS(\AV(\pi))=\WF(\pi)
$$
In other words the bijection (3)--(4) is induced by the Kostant--Sekiguchi correspondence.
\end{proposition}

@@

Consider the principal nilpotent element~$F_\pi$ defined in~\eqref{e:Fpi}, so by Lemma \ref{l:AV} $\AV(\pi)=K\cdot \psi(F_\pi)$. 
We use this as~$F_\theta$ in the previous subsection: define $E_\theta = \sum_{\alpha \in S(\lambda)}X_\alpha$ and $H_\theta  = [E_\theta, F_\theta]$; then $E_\theta \in \s$ and $(E_\theta, H_\theta, F_\theta)$ is an $\mathrm{SL}(2)$-triple. Define another $\mathrm{SL}(2)$-triple $(E_\R, F_\R, H_\R)$ by
\[ \begin{aligned}
  F_\R&=\frac12(-E_\theta+F_\theta-H_\theta)\\
  E_\R&=\frac12(E_\theta-F_\theta-H_\theta)\\
  H_\R&=E_\theta+F_\theta.
\end{aligned}\]
By the discussion in the previous section $\KS(K\cdot \psi(F_\theta))=\psi(\KS(K\cdot F_\theta))=\psi(G(\R)\cdot (-E_\R))=G(\R)\cdot \psi(-E_\R)$.
We need to show this equals $\WF(\pi)$. By Corollary \ref{cor:matumoto} it is enough to show $\Wh(\pi)=\w_{\psi(-E-\R)}$; by 
  by Lemma \ref{lem:Kost_and_AC} this is equivalent to $\K(\psi(-E_\R))$ meets $G(\R)\cdot\lambda$.

We compute $\K(\psi(-E_\R))=\psi(\K(-E_\R))\ni -E_\R-F_\R=H_\theta$ (by \eqref{def_F_R}).
Thus $\K(\psi(-E_\R))$ meets $\psi(H_\theta)$. 
Since  $\Psi(H_{\theta})$ and $\lambda$ are in the same Weyl chamber, by Proposition \ref{prop:AC_chamber} we conclude 
$\K(\psi(-E_\R))$ meets $\lambda$ as required.

@@

@@

Let's check Proposition \ref{Sekiguchi_result} in this case.
Suppose $\pi=\pi(\lambda_{ik})$. First assume $k>0$ and compute
$$
\begin{aligned}
\KS(\AV(\pi))&=  \KS(K\cdot E_\theta^*)\quad \text{by \eqref{e:AV}}\\
&=  \psi_c(\KS(K\cdot \psi_c\inv (E_\theta^*)))\\
  &=  \psi_c(\KS(K\cdot \frac 1c F_\theta))\\
&=\psi_C(G(\R)\cdot \frac 1c(-E_\R))\quad\text{(by \eqref{e:KS})}\\
&=  G(\R)\cdot \psi_c(\frac{-1}c F_\R))\\
&=G(\R)\cdot (-E_\R^*)\\
&=\WF(\pi)\quad \text{(by \ref{e:WF})}
\end{aligned}
$$
which agrees with 
Proposition \ref{Sekiguchi_result}.

Next, let's check Proposition \ref{p:Kostant}, 
which says $\Wh(\pi)=\w_X\Leftrightarrow \K(X)\cap\O_\pi\ne \emptyset$. 
Assume $k>0$. Then 
$$
\Wh(\pi(\lambda_{ik}))=G(\R)\cdot(-E_\R^*).
$$
So we compute
$$
\begin{aligned}
\K(-E_\R^*)&=\psi_c(\K(\psi_c\inv(-E_\R^*)))\\
&=\psi_c(\K( -\frac 1cF_\R))\\
&\ni \psi_c\frac 1c(-F_\R-E_\R)\\
&=\psi_c(\frac 1cH_\theta)=2H_\theta^*
\end{aligned}
$$
That is 
$$
\K(-E_\R^*)\cap G(\R)\cdot 2H_{\theta}^*\ne\emptyset
$$

>>>>

We need to pass to the dual Lie algebra $\g^*$, and it is important to do this as canonically as possible, i.e. independent of a choice
of isomorphism $\g\simeq\g^*$. First we define the nilpotent cone in $\g^*$ as follows. 
We say an element of $\g^*$ is nilpotent if its $G$-orbit is a (weak) cone, i.e. closed under multiplication by $\C^*$,
and \emph{regular} if $\dim_{\Cent(G)}(X)=\mathrm{rank}(G)$.
Let $\N^*$ be the set of nilpotent elements.
The principal nilpotent orbit $\O_p\in \N^*$ is the set of regular nilpotent elements, and $\overline{\O_p}=\N^*$.

The statements of the main results do not require the choice of an isomorphism $\g\simeq \g^*$. However such a choice appears in the proofs so we spell this out.
Having a $G(\R)$-invariant bilinear form on $\g(\R)$ is equivalent to having a $G(\R)$-equivariant isomorphism $\g(\R)\simeq \g^*(\R)$. 
Suppose we fix a non-degenerate, $G(\R)$-invariant bilinear form $\kappa$ on $\g(\R)$. This is equivalent to having a non-degenerate, $G(\C)$ and Galois invariant
bilinear form on $\g$. The restriction of the form each simple factor
is unique up to a non-zero constant. For example we can use the Killing form on $\gder$, extended arbitrarily by a non-degenerate form on the center.

@@

Using $\psi_\kappa$ we transfer the asymptotic cone, Kostant section, and Kostant Sekiguchi corresondence to $\g^*$. In each case these constructions are independent of the
choice of $\kappa$. See Section \ref{sec:Sekiguchi}.
In Section \ref{s:AV} we need to be more precise: we choose  $\kappa$ to be the Killing form. (More generally, we could take any non-degenerate
invariant form which is negative definite on the Lie algebra of a maximal compact subgroup.)

@@

It is well known that there is a natural bijection between nilpotent $G(\R)$-orbits on $\g(\R)$ and nilpotent $K$ orbits on $\s$. 
This is the Kostant--Sekiguchi
correspondence~\cite{sekiguchi}, which we write as $\KS(\O)$. It is a deep theorem of
Schmid and Vilonen (the main result of~\cite{SV1}) that the wavefront
set and the associated variety are related by this
correspondence. (As a historical aside, the Kostant--Sekiguchi
correspondence was discovered on the basis that the correspondence
$\AV(\pi) \leftrightarrow \WF(\pi)$ might be expressed in elementary
terms.)

@@

\begin{proposition} \label{Sekiguchi_result}
Suppose $\pi$ is a generic discrete series representation. Then
$$
\KS(\AV(\pi))=\WF(\pi)
$$
\end{proposition}

@@

Consider the principal nilpotent element~$F_\pi$ defined in~\eqref{e:Fpi}, so by Lemma \ref{l:AV} $\AV(\pi)=K\cdot \psi(F_\pi)$. 
We use this as~$F_\theta$ in the previous subsection: define $E_\theta = \sum_{\alpha \in S(\lambda)}X_\alpha$ and $H_\theta  = [E_\theta, F_\theta]$; then $E_\theta \in \s$ and $(E_\theta, H_\theta, F_\theta)$ is an $\mathrm{SL}(2)$-triple. Define another $\mathrm{SL}(2)$-triple $(E_\R, F_\R, H_\R)$ by
\[ \begin{aligned}
  F_\R&=\frac12(-E_\theta+F_\theta-H_\theta)\\
  E_\R&=\frac12(E_\theta-F_\theta-H_\theta)\\
  H_\R&=E_\theta+F_\theta.
\end{aligned}\]
By the discussion in the previous section $\KS(K\cdot \psi(F_\theta))=\psi(\KS(K\cdot F_\theta))=\psi(G(\R)\cdot (-E_\R))=G(\R)\cdot \psi(-E_\R)$.
We need to show this equals $\WF(\pi)$. By Corollary \ref{cor:matumoto} it is enough to show $\Wh(\pi)=\w_{\psi(-E_
  \R)}$; by 
  by Lemma \ref{lem:Kost_and_AC} this is equivalent to $\K(\psi(-E_\R))$ meets $G(\R)\cdot\lambda$.

We compute $\K(\psi(-E_\R))=\psi(\K(-E_\R))\ni -E_\R-F_\R=H_\theta$ (by \eqref{def_F_R}).
Thus $\K(\psi(-E_\R))$ meets $\psi(H_\theta)$. 
Since  $\psi(H_{\theta})$ and $\lambda$ are in the same Weyl chamber, by Proposition \ref{prop:AC_chamber} we conclude 
$\K(\psi(-E_\R))$ meets $\lambda$ as required.

@@

We illustrate these results with the illustrative example of $SL(2,\R)$,.

@@

Let's check Proposition \ref{Sekiguchi_result} in this case, which says $\KS(\AV(\pi))=\WF(\pi)$.
Suppose $\pi=\pi(\lambda_{ik})$. First assume $k>0$ and compute
$$
\begin{aligned}
\KS(\AV(\pi))&=  \KS(K\cdot E_\theta^*)\quad \text{by \eqref{e:AV}}\\
&=  \psi_c(\KS(K\cdot \psi_c\inv (E_\theta^*)))\\
  &=  \psi_c(\KS(K\cdot \frac 1c F_\theta))\\
&=\psi_c(G(\R)\cdot \frac 1c(-E_\R))\quad\text{(by \eqref{e:KS})}\\
&=  G(\R)\cdot \psi_c(\frac{-1}c F_\R))\\
&=G(\R)\cdot (-E_\R^*)\\
&=\WF(\pi)\quad \text{(by \ref{e:WF})}
\end{aligned}
$$
which agrees with 
Proposition \ref{Sekiguchi_result}. If $k<0$ replace $E_\theta^*$ with $F_\theta^*$, and conclude $\KS(\AV(\pi))=G(\R)\cdot E_\R^*$, in accordance with
Proposition \ref{Sekiguchi_result}.

Next, let's check Proposition \ref{p:Kostant}, 
which says $\Wh(\pi)=\w_X\Leftrightarrow \K(X)\cap\O_\pi\ne \emptyset$. 
Assume $k>0$. Then 
$$
\Wh(\pi(\lambda_{ik}))=G(\R)\cdot(-E_\R^*).
$$
So we compute
$$
\begin{aligned}
\K(-E_\R^*)&=\psi_c(\K(\psi_c\inv(-E_\R^*)))\\
&=\psi_c(\K( -\frac 1cF_\R))\\
&\ni \psi_c\frac 1c(-F_\R-E_\R)\\
&=\psi_c(\frac 1cH_\theta)=2H_\theta^*
\end{aligned}
$$
That is 
$$
\K(-E_\R^*)\cap G(\R)\cdot 2H_{\theta}^*\ne\emptyset\quad (k<0).
$$
The case $k<0$ is similar; we conclude
$\Wh(\pi_{\lambda_ik})=G(\R)E_\R^*$, and $\K(E_\R^*)\cap G(\R)\cdot (-2H_\theta)\ne \emptyset$.
